---
title: "Results And Discussion"
author: "Paul Hegedus"
output: html_document
---
This page documents the analysis process after sensitivity analysis has been performed in each field. This includes the importing of sensitivity analysis outputs, aggregation into one dataset, and evaluation of patterns between fields and years. Discussion is provided on each result as it pertains to the project.

```{r, message = FALSE, warning = FALSE}
library(magrittr)
library(ggplot2) 
source("R/importCleanSelectFuns.R")
```

## Import Data
For each field that sensitivity analysis was performed, the resulting tables need to be imported and amended with field specific information before combining. 

```{r}
fields <- c("carlinwest", "henrys", "minnies",
            "sec1east", "sec1west", "sec35mid",
            "sre1314") 
fields <- as.list(fields) %>%
  `names<-`(fields)

for (i in 1:length(fields)) {
  fldFiles <- list.files(paste0("results/", fields[i]))
  fldFiles <- fldFiles[grep("sensTableCent", fldFiles)]
  
  fldList <- rep(list(NA), length(fldFiles))
  yrs <- rep(list(NA), length(fldFiles))
  for (j in 1:length(fldFiles)) {
    yrs[[j]] <- stringr::str_locate(fldFiles[j], "20")
    yrs[[j]] <- stringr::str_sub(fldFiles[j], yrs[[j]][1,1], yrs[[j]][1,1] + 3)
    fldList[[j]] <- impDat(fldFiles[j], paste0("results/", fields[i]))
    fldList[[j]]$year <- yrs[[j]]
  }
  fields[[i]] <- do.call(rbind, fldList)
}
dat <- do.call(rbind, fields)
```


## Analysis
Since we are interested in the most influential parameters on each response variable, we subset out only the most elastic (Elasticity.value == 1) for each field and year and response variable.

```{r}
fields <- c("carlinwest", "henrys", "minnies",
            "sec1east", "sec1west", "sec35mid",
            "sre1314") 
fields <- as.list(fields) %>%
  `names<-`(fields)

elasticOnes <- subset(dat, dat$Elasticity.value == 1)
elasticOnes$Parameter <- stringr::str_replace(elasticOnes$Parameter, "cent_", "")
for (i in 1:length(fields)) {
  fields[[i]] <- subset(elasticOnes, elasticOnes$field == fields[[i]])
}
elasticOnes
```

Below, a barplot of most influential parameter across all fields and years and responses is shown. 'aa_n' = As-applied nitrogen fertilizer (lbs N/acre), 'aspect_rad' = aspect in radians, 'elev' = elevation in meters, 'veg_2py' = NDVI for January 1st - December 31st from two years prior, 'veg_cy' = NDVI from January 1st - March 30th of the year of harvest, and 'veg_py' =  NDVI for January 1st - December 31st from the previous year.

```{r}
ggplot(elasticOnes) +
  geom_bar( aes(Parameter, fill = ResponseVar)) +
  theme_bw() + 
  scale_y_continuous(name = "Frequency", limits = c(0, 20)) +
  scale_fill_manual(name = "Respone Variable", 
                    labels = c("Net-Return", "Yield", "Protein"), 
                    values = c("#7CAE00", "#F8666D", "#00BFC4")) +
  ggtitle("Most Elastic Parameter",
          subtitle = "n = 42")
ggsave(plot = last_plot(),
       filename = paste0("results/elasticParms.png"),
       width = 7.5, height = 7.5, units = "in")
```

## Results & Discussion
Results in the following paragraph relate to the barplot above. The most common parameter that had the largest influence on any of the response variables was elevation, followed by as-applied nitrogen. Elevation was the most influential parameter for both net-return and yield across all fields and years, however elevation and nitrogen fertilzier were equally influential on protein. This indicates that although fertilizer is the input that farmers have the most control over and is, of course, a plant macronutrient, is not the biggest driver of yield, protein, and net-return in all cases. Vegetation indices (NDVI) from previous years also had an influence on responses, albeit in less instances. The amount of 'greenness' reflected by a crop, measured as NDVI, relates to the productivity of the plant in that timeframe, so these results indicate that the amount of crop productivity in previous years is having an influence on the productivity and quality of the crop in the current year. 

The rest of these results reference the table of the most elastic parameters above. For a given field and year combination, the most elastic parameter was the same for yield, protein, and net-return in 6 out of 14 instances. This shows that about half the time, the same driver is the most influential on each response, a result that is not entirely surprising due to the relationship between yield, protein, and net-return. However, as this instance only occurs half the time, this indicates that not only is there field and time specificity, there is specificity in management for different responses, even in the same field and year.

Related, and of note, is that in 12 out of 14 instances, the most elastic covariate for yield and net-return were the same. This highlights the tight relationship between yield and net-return, which was also seen in the correlation matrices of each field and year, where $R^2$ values were observed in the range of 0.9 to 0.95 between yield and net-return. However, the caveat must be stated that farmers do not know the price they will receive in the upcoming year, and that this is a source of uncertainty in net-return not addressed in this work. 

When comparing the same response variable in the same field across years, it was observed that the same covariate had the most influence on the response variable in both years in 5 out of 21 cases. This shows that there 

## Conclusions
Understanding the primary sources of field-specific variation in production and quality across the field and over time enables producers to identify if, when, and where nitrogen fertilizer should be applied. Equally important, the spatial and temporal resolution at which variability in agroecosystem performance can be detected creates a mechanism to monitor and ground-truth ecosystem response to global change.

This project demonstrates the importance of field specificity in management of small grain agroecosystems in Montana. In ~50% of cases, the same covariate was the most influential on yield, protein, and net-return, otherwise, different factors driving each response, meaning that different drivers for different response variables even in the same field and year. Even in fields bordering each other, different covariates had the most influence on each response variable in each year (i.e. sec1east & sec1west), indicating that there are different drivers for responses across fields in the same year (even on same farm). In ~25% of cases, the same covariate was the most elastic across years, ~75% of the time, different covariate was most influential on the response in the same field in different years, meaning that different drivers most influential on responses across years in the same field.












